### 2020-02-20 Thesis-Chapter-Background.tex
Begin writing introduction

\chapter{Background}

\section{Arguments For and Against}

\subsection{Arguments For}

\subsection{Arguments Against}

\section{Technical State of the Art}

\subsection{Historical Approaches}

\subsection{Modern Proposals}

\section{Paths Ahead}

\subsection{Legislative Priorities}

\subsection{Technical Research}