### 2020-02-20 Thesis-Chapter-Intro.tex

\chapter{Introduction}

Exceptional access is a wicked problem. Characterized by a dynamic technological environment, disagreement about
underlying values, and resistance to a clear solution, the debate on exceptional access refuses to end. This paper will
not attempt to end the debate, but to inform it with additional structure.

\section{Motivation}

\begin{itemize}
    \item In the pursuit of data privacy, legislative action is part of the threat model
\end{itemize}

\section{Premises}

\begin{itemize}
    \item The lack of convincing of legislative bodies and law enforcement agencies demands that the technical community continue the discussion
    \item The ability to access digital data is important and will be increasingly important for investigations
    \item Various forms of encryption are foundational all elements of online security
    \item Mass surveillance is a grave danger that representative governments are not immune from
    \item International and human rights considerations significantly constrain workable solutions
    \item Transparency to data access is critical for public trust
\end{itemize}

\section{Contribution}

\begin{itemize}
    \item Argument maps formalizing the debate
    \item A broad threat model describing the environment
    \item A structure for organizing future progress
\end{itemize}