### 2020-02-26 Thesis-Chapter-Background.tex
Work on intro/background

\chapter{Background}

Private use of encryption has a contentious history. [Some highlights.]

Electronic encryption must be understood as a technological development embedded in the longer history of privacy and
human rights. [This is where I'd like to understand more of the legal side. With a historical perspective.]

Privacy is not encryption's only purpose, however. Encryption is the technical foundation for many forms of computer and
network security. [Properties: confidentiality, authentication, integrity, non-repudiation; primitives: hashed
passwords, digital signatures, ...; technologies: PKI, TLS, SSO/Kerberos, VPNs, disk encryption, blockchains (might get
eye-rolls).]

Electronic encryption's dual role as enabler of perfect privacy and cornerstone of computer security may be the central
conflict in the exceptional access debate. Those who focus on its role in privacy object to its absolute power. Those
who focus on its role in security rely on its absolute power.

% Encryption requires absolute cryptographic fidelity in order to provide security.

\section{Arguments For and Against}

\subsection{Arguments For}

\subsection{Arguments Against}

\section{Technical State of the Art}

\subsection{Historical Approaches}

\subsection{Modern Proposals}

\section{Paths Ahead}

\subsection{Legislative Priorities}

\subsection{Technical Research}