### 2020-02-28 Thesis-Chapter4-ThreatModel.tex
Improve file naming scheme, work on background

\chapter{Threat Model}
\label{chap-threatmodel}

In this chapter, develop a threat model using data flow diagrams. May need separate ones for different scenarios (data
at rest on encrypted devices, data in motion using E2E), but want at least one high level diagram encapsulating
everything.