### 2020-03-28 Thesis-Chapter5-PathsAhead.tex
Add reading sources; update background

\chapter{Paths Ahead}
\label{chap-pathsahead}

In this chapter, describe actions to take in discussion, regulation, and research.

\section{Legislative Priorities}

\begin{itemize}
    \item Reject EA until technical capabilities are ready \cite{varia_2018}
    \item Be forthcoming with data so that we know what to work on \cite{devlin_2018}
    \item Don't prescribe technical methods, but be clear about principles \cite{matyas_incommensurability_2018}
        (e.g., ECPA became outdated because was too tied to specific technology \cite{shamsi_2011})
    \item Fund the research \cite{varia_2018}
\end{itemize}

\section{Technical Research}

\begin{itemize}
    \item Acknowledge the moral character of cryptography \cite{rogaway_moral_2015}
    \item Do the research
    \item Focus on principles \cite{levy_robinson_2018}
    \item Be clear on definitions \cite{varia_2018}, but don't expect govt to provide those (unlike
            \cite{abelson_2015}), and clearly define their interfaces with principles
            \cite{matyas_incommensurability_2018}
\end{itemize}