### 2020-04-05 Thesis-Glossary.tex
Work on background; rearrange sections a bit

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                       GLOSSARY
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Reference using the schema
%
%   \ac{}   -- general: abbreviation except first time is full
%   \acs{}  -- short: abbreviation
%   \acl{}  -- long: definition
%   \acf{}  -- full: definition (abbreviation)
%
%   pluralization       -- append "p", e.g. \acp{}
%   capitalize first    -- capitalize first of command, e.g. \Ac{}
%   capitalize all      -- capitalize all of command, e.g. \AC{}
%
%   See page 85 of the documentation:
%   http://ctan.math.utah.edu/ctan/tex-archive/macros/latex/contrib/glossaries-extra/glossaries-extra-manual.pdf

\makeglossaries

\newcommand{\newterm}[3]{
    \newglossaryentry{#1}{
        name={#2},
        description={#3}
    }
}
\newcommand{\term}{\ac}
\newcommand{\Term}{\Ac}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                          TERMS

% NOTE: Do not end with a period. The package adds this for you.
% IDEA: auto-update glossary.md. This should be the master, because it is more fully specified

% Start terms
\newterm{clipperchip}{Clipper Chip}{
    A 1993 initiative by the Clinton administration to provide ``the public with strong cryptographic tools without
    sacrificing the ability of law enforcement and intelligence agencies to access unencrypted versions of those
    communications'' via hardware additions to consumer electronics. The proposal faced opposition, and died in 1994
    when security researcher Matt Blaze discovered flaws that allowed users to subvert the Clipper Chip mechanisms
}

\newterm{PRISM}{PRISM}{
    An NSA program unveiled by Edward Snowden involving the mass collection of  internet communications data of non-U.S.
    persons and those communicating with them
}

\newterm{horsemen}{the Four Horsemen}{
    Four reasons that governments and law enforcement agencies use to undercut public support for strong encryption. The
    four horsemen are terrorists, pedophiles (child pornographers), drug dealers (and traffickers), and money launderers
    (or kidnappers)
}
\newterm{horsemen-full}{The Four Horsemen of the Infocalypse}{The long name for ``The Four Horsemen''}

\newterm{plaintext}{plaintext}{Unencrypted data}
\newterm{ciphertext}{ciphertext}{Encrypted data}

\newterm{cryptography}{cryptography}{
    The study of techniques for communicating secretly in the presence of third parties
}

\newterm{cipher}{cipher}{
    A tool or algorithm to perform encryption, translating between plaintext and ciphertext
}
\newterm{key}{key}{Secret information required to operate the cipher and perform encryption and decryption}

\newterm{encryption}{encryption}{The process of translating from plaintext to ciphertext}
\newterm{decryption}{decryption}{The process of translating from ciphertext to plaintext}

\newterm{public-key-crypto}{public key cryptography}{
    Another name for asymmetric encryption
}
\newterm{asym-crypto}{asymmetric cryptography}{
    A cipher scheme where encryption and decryption are performed with separate, paired keys, called the public key and
    the private key
}
\newterm{sym-crypto}{symmetric cryptography}{
    A cipher scheme where encryption and decryption are performed with the same key
}
% End terms

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                       ACRONYMS

% Start acronyms
\newabbreviation{AECA}{AECA}{Arms Export Control Act}
\newabbreviation{AV}{AV}{anti-virus}
\newabbreviation{AWA}{AWA}{All Writs Act}
\newabbreviation{CALEA}{CALEA}{Communications Assistance for Law Enforcement Act}
\newabbreviation{CFAA}{CFAA}{Computer Fraud and Abuse Act}
\newabbreviation{CP}{CP}{child pornography}
\newabbreviation{CSAM}{CSAM}{child sexual abuse material}
\newabbreviation{DE}{DE}{device encryption}
\newabbreviation{DIM}{DIM}{data in motion}
\newabbreviation{DAR}{DAR}{data at rest}
\newabbreviation{E2E}{E2E}{end to end}
\newabbreviation{EA}{EA}{exceptional access}
\newabbreviation{ECPA}{ECPA}{Electronic Communications Privacy Act}
\newabbreviation{ED}{ED}{ephemeral data}
\newabbreviation{FBI}{FBI}{Federal Bureau of Investigation}
\newabbreviation{FISA}{FISA}{Foreign Intelligence Surveillance Act}
\newabbreviation{FISAAA}{FISAAA}{FISA Amendments Act}
\newabbreviation{FISC}{FISC}{Foreign Intelligence Surveillance Court}
\newabbreviation{ITAR}{ITAR}{International Traffic in Arms Regulations}
\newabbreviation{KEK}{KEK}{key encrypting key}
\newabbreviation{LE}{LE}{law enforcement}
\newabbreviation{LEA}{LEA}{law enforcement agency}
\newabbreviation{LEO}{LEO}{law enforcement officer}
\newabbreviation{NED}{NED}{non-ephemeral data}
\newabbreviation{NSA}{NSA}{National Security Agency}
\newabbreviation{NSL}{NSL}{National Security Letter}
\newabbreviation{PAA}{PAA}{Protect America Act}
\newabbreviation{PGP}{PGP}{Pretty Good Privacy}
\newabbreviation{PIN}{PIN}{personal identification number}
\newabbreviation{PKI}{PKI}{public key infrastructure}
\newabbreviation{PW}{PW}{password}
\newabbreviation{SE}{SE}{strong encryption}
\newabbreviation{SSO}{SSO}{single sign-on}
\newabbreviation{TLS}{TLS}{Transport Layer Security}
\newabbreviation{VPN}{VPN}{virtual private network}
% End acronyms

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                    UN-OFFICIAL

\newcommand{\US}{U.S. } % Not good; doesn't handle spacing, periods at ends of sentences very well