### 2020-09-05 Thesis-Chapter4-ThreatModel.tex
Improve argdown map generation build process

\chapter{Threat Model}
\label{chap-threatmodel}

In this chapter, develop a threat model using data flow diagrams. May need separate ones for different scenarios (data
at rest on encrypted devices, data in motion using E2E), but want at least one high level diagram encapsulating
everything.

% For analysis of threat actors, see bindenagel_discussion_2017

% Generalize for DAR and DIM, then apply to specific examples. This proves that my analysis is helpful. Ex: apply to email: is a DIM _and_ DAR problem due to the server.