### 2020-09-26 Thesis-Abstract.tex
Update abstract and TODOs

Public debate has resumed over whether encryption systems should support alternative means of decryption intended for
law enforcement use, called exceptional access (EA). Rather than a renege on a solemn promise made at the end of the
1990s ``crypto war,'' this represents a valid reassessment of optimal policy in light of changing circumstances.
Achieving proper balance between privacy and access in the context of constantly changing society and technology is a
wicked problem that has and will evade a permanent solution. As policymakers consider next steps, it behooves the
technical community to stay engaged. Although the introduction of EA would inevitably introduce risk, the quality of the
technical and regulatory approach can make a substantial difference. Furthermore, if one considers ham-fisted
legislative action and malicious abuse of cryptosystems as part of the threat model, well-designed EA may reduce overall
risk.

The root of the conflict lies in cryptography's dual role as enabler of radical privacy and cornerstone of security. The
emergence of strong encryption triggered the first crypto war, and its proliferation is causing the second. In the face
of both polarized and conciliatory voices, I analyze strategies that do and do not work on wicked problems and promote
an iterative approach to the problem of encryption and EA. Along the way I illustrate the components of the debate in
argument maps and demonstrate the security risks with data flow diagrams, making clear distinctions for the separate
problems of EA for data in motion and data at rest.