### 2020-11-15 Thesis-Chapter6-PathsAhead.tex
Clean up conclusion notes

\chapter{Paths Ahead}
\label{chap-pathsahead}

% We have a wicked problem
% I propose we handle it a certain way
% Part of this is cooperative debate, which I tried to capture in argument maps
% Another element is focusing on specific proposals, so we dove into LDAWMSR

%% Prof. Walsh:
%% The dynamic regulatory environment and competing governmental branches at the executive, legislative, and judicial
%% level will continue to shape the...

% What can we conclude?
%   - It may not feel like it, but (meaningful) debate and discussion counts as progress for a wicked problem
%   - Neither side has the whole picture, but the "no EA" arguments are currently stronger
%   - A threat model including malicious users and threatening governments makes EA make more sense
%   - Device escrow with secure hardware represents true progress towards acceptably secure EA, so hope is not lost
% Where do we go from here?
%   - This was an attempt an "OO" iteration in the OODA; we need more, with involvement all around
%   - committee_decrypting_2018 presented a framework to use to judge proposals. I have a simple version of it.
%     here are where we are weak:
%       - X
%       - Y
%       - Z
%   - Government:
%       - Take the low hanging fruit first, please
%       - Fund research, don't mislead, don't be premature
%           - If you're going to compare it to going to the moon, fund us like it!
%       - Take on weak area XYZ
%   - Tech:
%       - Do research, provide clear security advice, be vigilant
%       - Take on weak area XYZ
%   - Either:
%       - Know that there is no stable solution, balance rules and principles, serve your neighbor
% see varia_2018 (maybe a good quote?)

In this chapter, describe actions to take in discussion, regulation, and research.

\section{Legislative Priorities}

\begin{itemize}
    \item Reject EA until technical capabilities are ready \cite{varia_2018}
    \item Be forthcoming with data so that we know what to work on \cite{devlin_2018}
    \item Don't prescribe technical methods, but be clear about principles \cite{matyas_incommensurability_2018}
        (e.g., ECPA became outdated because was too tied to specific technology \cite{shamsi_2011})
    \item Fund the research \cite{varia_2018} (like \cite{goss_hr2616_1999})
\end{itemize}

\section{Technical Research}

\begin{itemize}
    \item Acknowledge the moral character of cryptography \cite{rogaway_moral_2015}
    \item Do the research
    \item Focus on principles \cite{levy_robinson_2018}
    \item Be clear on definitions \cite{varia_2018}, but don't expect govt to provide those (unlike
            \cite{abelson_2015}), and clearly define their interfaces with principles
            \cite{matyas_incommensurability_2018}
\end{itemize}