### 2020-11-20 Thesis-Glossary.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                       GLOSSARY
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Reference using the schema
%
%   \ac{}   -- general: abbreviation except first time is full
%   \acs{}  -- short: abbreviation
%   \acl{}  -- long: definition
%   \acf{}  -- full: definition (abbreviation)
%
%   pluralization       -- append "p", e.g. \acp{}
%   capitalize first    -- capitalize first of command, e.g. \Ac{}
%   capitalize all      -- capitalize all of command, e.g. \AC{}
%
%   See page 85 of the documentation:
%   http://ctan.math.utah.edu/ctan/tex-archive/macros/latex/contrib/glossaries-extra/glossaries-extra-manual.pdf

\makeglossaries

\newcommand{\newterm}[3]{
    \newglossaryentry{#1}{
        name={#2},
        description={#3}
    }
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                          TERMS

% NOTE: Do not end with a period. The package adds this for you.

% Start terms
\newterm{wicked-prob}{wicked problem}{
    A problem lacking clear formulations, causes, resolutions, or metrics. Consistenly difficult to solve, and often
    social or political in nature
}

\newterm{the-cw1}{the first crypto war}{
    The flurry of evolution and conflict in encryption policy during the 1990s, characterized by episodes around export
    controls, \ac{PGP}, the \ac{clipperchip}, and early key escrow proposals
}
\newterm{the-cw2}{the second crypto war}{
    The flurry of conflict in encryption policy during the late 2010s to present, characterized by angst from the
    Snowden revelations, the proliferation of strong encryption, and the Apple vs. FBI case
}

\newterm{clipperchip}{Clipper Chip}{
    A 1993 initiative by the Clinton administration to provide ``the public with strong cryptographic tools without
    sacrificing the ability of law enforcement and intelligence agencies to access unencrypted versions of those
    communications'' via hardware additions to consumer electronics. The proposal faced opposition, and died in 1994
    when security researcher Matt Blaze discovered flaws that allowed users to subvert the Clipper Chip mechanisms
}

\newterm{PRISM}{PRISM}{
    An NSA program unveiled by Edward Snowden involving the mass collection of  internet communications data of non-U.S.
    persons and those communicating with them
}

\newterm{horsemen}{Four Horsemen}{
    Four reasons that governments and law enforcement agencies use to undercut public support for strong encryption. The
    four horsemen are terrorists, pedophiles (child pornographers), drug dealers (and traffickers), and money launderers
    (or kidnappers)
}
\newterm{horsemen-full}{the Four Horsemen of the Infocalypse}{The long name for ``The Four Horsemen''}

\newterm{plaintext}{plaintext}{Unencrypted data}
\newterm{ciphertext}{ciphertext}{Encrypted data}

\newterm{cryptography}{cryptography}{
    The study of techniques for communicating secretly in the presence of third parties
}

\newterm{cipher}{cipher}{
    A tool or algorithm to perform encryption, translating between plaintext and ciphertext
}
\newterm{key}{key}{Secret information required to operate the cipher and perform encryption and decryption}

\newterm{encryption}{encryption}{The process of translating from plaintext to ciphertext}
\newterm{decryption}{decryption}{The process of translating from ciphertext to plaintext}

\newterm{public-key-crypto}{public key cryptography}{
    Another name for asymmetric encryption
}
\newterm{asym-crypto}{asymmetric cryptography}{
    A cipher scheme where encryption and decryption are performed with separate, paired keys, called the public key and
    the private key
}
\newterm{sym-crypto}{symmetric cryptography}{
    A cipher scheme where encryption and decryption are performed with the same key
}

\newterm{forward-secrecy}{forward secrecy}{
    A property of \ac{DIM} encryption protocols that ensures that a leaked private key or session key does not
    compromise any other private key or session key
}
\newterm{replay-protection}{replay protection}{
    A property of \ac{DIM} encryption protocols that ensures that messages cannot be replayed by an attacker without
    detection
}

\newterm{disk-encryption}{disk encryption}{
    An application of ecnryption for \ac{DAR} that involves encrypting the contents
}

\newterm{snowden}{Snowden}{
    Edward Snowden was an CIA contractor who leaked large volumes of data on NSA programs such as PRISM that performed
    mass collection of internet communications for surveillance
}

\newterm{lawful-hacking}{lawful hacking}{
    An approach to digital evidence and intelligence gathering that involves deliberate computer and network
    manipulation which is ordinarily illegal but has been cleared for law enforcement and intelligence purposes
}

\newterm{key-escrow}{key escrow}{
    An approach to \ac{EA} that relies on storing additional copies of the encryption key or storing information that
    can be used to derive additional additional copies
}

\newterm{classical-method}{classical analytic method}{
    A policy making strategy that emphasizes rationalism and functions by setting goals, identifying problems,
    evaluating alternatives, implementing solutions, and analyzing outcomes in order to correct errors
}

\newterm{incrementalism}{incrementalism}{
    A policy making strategy that functions by taking successive steps based on the status quo and chosen through
    comparative analysis; also known as ``muddling through''
}

\newterm{muddling-through}{muddling through}{
    See ``incrementalism''
}

\newterm{argmap}{argument map}{
    A graphical representation of the logical structures and relationships between statements, premises, and
    conclusions, used to disentangle complex arguments used in wicked problems
}

\newterm{LDAWMSR}{LDAWMSR}{
    A reference to computer scientist and security researcher Stefan Savage's device key escrow proposal, ``Lawful
    Device Access without Mass Surveillance Risk: A Technical Design Discussion'' \cite{savage_lawful_2018}
}
% End terms

\newcommand{\ldawmsR}[0]{\ii{\ac{LDAWMSR}}}
\newcommand{\ldawmsr}[0]{\ldawmsR~}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                       ACRONYMS

% Start acronyms
\newabbreviation{AECA}{AECA}{Arms Export Control Act}
\newabbreviation{APT}{APT}{advanced persistent threat}
\newabbreviation{AV}{AV}{anti-virus}
\newabbreviation{AWA}{AWA}{All Writs Act}
\newabbreviation{CALEA}{CALEA}{Communications Assistance for Law Enforcement Act}
\newabbreviation{CESA}{CESA}{Cyberspace Electronic Security Act}
\newabbreviation{CFAA}{CFAA}{Computer Fraud and Abuse Act}
\newabbreviation{CIA}{CIA}{Central Intelligence Agency}
\newabbreviation{CP}{CP}{child pornography}
\newabbreviation{CSAM}{CSAM}{child sexual abuse material}
\newabbreviation{CSAV}{CSAV}{computer supported argument visualization}
\newabbreviation{DAR}{DAR}{data at rest}
\newabbreviation{DE}{DE}{device encryption}
\newabbreviation{DFD}{DFD}{data flow diagram}
\newabbreviation{DIM}{DIM}{data in motion}
\newabbreviation{DOS}{DOS}{denial of service}
\newabbreviation{E2E}{E2E}{end to end}
\newabbreviation{E2EE}{E2EE}{end to end encryption}
\newabbreviation{EA}{EA}{exceptional access}
\newabbreviation{ECPA}{ECPA}{Electronic Communications Privacy Act}
\newabbreviation{ED}{ED}{ephemeral data}
\newabbreviation{FBI}{FBI}{Federal Bureau of Investigation}
\newabbreviation{FISA}{FISA}{Foreign Intelligence Surveillance Act}
\newabbreviation{FISAAA}{FISAAA}{FISA Amendments Act}
\newabbreviation{FISC}{FISC}{Foreign Intelligence Surveillance Court}
\newabbreviation{GCHQ}{GCHQ}{(UK) Government Communications Headquarters}
\newabbreviation{HSM}{HSM}{Hardware Security Module}
\newabbreviation{IBIS}{IBIS}{issue-based information system}
\newabbreviation{ITAR}{ITAR}{International Traffic in Arms Regulations}
\newabbreviation{KEK}{KEK}{key encrypting key}
\newabbreviation{LE}{LE}{law enforcement}
\newabbreviation{LEA}{LEA}{law enforcement agency}
\newabbreviation{LEO}{LEO}{law enforcement officer}
\newabbreviation{NED}{NED}{non-ephemeral data}
\newabbreviation{NSA}{NSA}{National Security Agency}
\newabbreviation{NSL}{NSL}{National Security Letter}
\newabbreviation{PAA}{PAA}{Protect America Act}
\newabbreviation{PGP}{PGP}{Pretty Good Privacy}
\newabbreviation{PIN}{PIN}{personal identification number}
\newabbreviation{PKI}{PKI}{public key infrastructure}
\newabbreviation{PW}{PW}{password}
\newabbreviation{SE}{SE}{strong encryption}
\newabbreviation{SSO}{SSO}{single sign-on}
\newabbreviation{TLS}{TLS}{Transport Layer Security}
\newabbreviation{UID}{UID}{Unique ID}
\newabbreviation{VPN}{VPN}{virtual private network}
% End acronyms

% Custom acronyms

\newabbreviation{USAPATRIOTAct}{USA PATRIOT Act}{Uniting and Strengthening America by Providing Appropriate Tools Required to Intercept and Obstruct Terrorism Act}
\newabbreviation{USAFREEDOMAct}{USA FREEDOM Act}{Uniting and Strengthening America by Fulfilling Rights and Ensuring Effective Discipline Over Monitoring Act}
\newabbreviation{SAFEAct}{SAFE Act}{Security and Freedom Through Encryption Act}
\newabbreviation{PROCODEAct}{Pro-C0DE Act}{Promotion of Commerce On-Line in the Digital Era Act}
\newabbreviation{ENCRYPTAct}{ENCRYPT Act}{Ensuring National Constitutional Rights for Your Private Telecommunications Act}
\newabbreviation{EARNITAct}{EARN IT Act}{Eliminating Abusive and Rampant Neglect of Interactive Technologies Act}
\newabbreviation{LEADAct}{LEAD Act}{Lawful Access to Encrypted Data Act}
\newabbreviation{OODAloop}{OODA Loop}{\ii{Observe-Orient-Decide-Act} Loop}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                    UN-OFFICIAL

\newcommand{\US}{U.S. } % Not good; doesn't handle spacing, periods at ends of sentences very well