### 2020-12-10 Thesis-Abstract.tex
Editorial improvements

Public debate has resumed on the topic of exceptional access (EA), which is alternative means of decryption intended for
law enforcement use. The resumption of this debate is not a renege on a resolute promise made at the end of the 1990s
``crypto war''; rather, it represents a valid reassessment of optimal policy in light of changing circumstances. The
imbalance between privacy, access, and security in the context of constantly changing society and technology is a wicked
problem that has and will continue to evade a permanent solution. As policymakers consider next steps, it is necessary
that the technical community to remain engaged. Although the introduction of EA would inevitably add risk, the quality
of the technical and regulatory approach can make a substantial difference. Furthermore, if one considers hard-line
legislative action and malicious abuse of cryptosystems as part of the threat model, well-designed EA may reduce risk
overall.

The root of the conflict lies in cryptography's dual role as an enabler of unprecedented privacy and a cornerstone of
security. The emergence of strong encryption incited the first crypto war, and its proliferation is causing the second.
In response to both polarized and conciliatory voices, this paper analyzes strategies for confronting wicked problems
and proposes an iterative approach to the case of encryption and EA. Along the way, it illustrates the components of the
debate in argument maps and demonstrate the security risks with data flow diagrams and threat analysis, focusing on one
EA proposal in particular, Stefan Savage's ``Lawful Device Access without Mass Surveillance Risk.''